---
title: "simpleRegress"
author: "Emann Malik"
date: "April 8, 2021"
output: 
  html_document:
    number_sections: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
  
The LENGTH ~ NECK model has a residual standard error (the standard deviation of the plotted points from the normal curve) of 5.376 and an adjusted R-squared (the percent of mean model error explained) of 0.7476.
  
## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
  
A bear with a neck of 17 is predicted to have a length of about 52.77.  
A bear with a neck of 22 is predicted to have a length of about 60.99.  
A bear with a neck of 26 is predicted to have a length of about 67.58.  
  
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).
  
### LENGTH ~ AGE 
  
```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)
```
  
The LENGTH ~ AGE model has a residual standard error of 7.511 and an adjusted R-squared of 0.5073. 
  
Compared to the LENGTH ~ NECK model, this one has a higher residual standard error and a lower adjusted R-squared.  
  
### LENGTH ~ HEADLEN
  
```{r}
HeadLenModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(HeadLenModel)
summary.lm(HeadLenModel)
```
  
The LENGTH ~ HEADLEN model has a residual standard error of 4.235 and an adjusted R-squared of 0.8434. 
  
Compared to the LENGTH ~ NECK model, this one has a lower residual standard error and a higher adjusted R-squared.    
  
### LENGTH ~ HEADWTH  
  
```{r}
HeadWthModel <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(HeadWthModel)
summary.lm(HeadWthModel)
```  
  
The LENGTH ~ HEADWTH model has a residual standard error of 7.32 and an adjusted R-squared of 0.5321. 
  
Compared to the LENGTH ~ NECK model, this one has a higher residual standard error and a lower adjusted R-squared.   
  
### LENGTH ~ CHEST  
  
```{r}
chestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestModel)
summary.lm(chestModel)
```   
  
The LENGTH ~ CHEST model has a residual standard error of 4.939 and an adjusted R-squared of 0.7869. 
  
Compared to the LENGTH ~ NECK model, this one has a lower residual standard error and a higher adjusted R-squared.   
  
### LENGTH ~ WEIGHT  
  
```{r}
weightModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weightModel)
summary.lm(weightModel)
```  
  
The LENGTH ~ WEIGHT model has a residual standard error of 5.432 and an adjusted R-squared of 0.7423. 
  
Compared to the LENGTH ~ NECK model, this one has a higher residual standard error and a lower adjusted R-squared.   
  
Both these models have a residual standard error and an adjusted R-squared that are very close to one another.
  
## Summary Statement
  
We created five models using five variables (AGE, HEADLEN, HEADWTH, CHEST, and WEIGHT) from the bears data set to find out what, if any, variables are better than NECK to predict the LENGTH of a bear.   
  
The best models (with a lower residual standard error and higher adjusted R-squared than that of LENGTH ~ NECK) are LENGTH ~ HEADLEN and LENGTH ~ CHEST. 
  
Out of the two, LENGTH ~ HEADLEN has the better values (the points are the closest to the normal curve and there is the highest percent of mean model error explained of all the other models) and best fit. 
  
The next best model for predicting is LENGTH ~ CHEST before LENGTH ~ NECK comes in third place. The rest of the models had a higher residual standard error and a lower adjusted R-squared than LENGTH ~ NECK.
  