---
title: "simpleRegress"
author: "Emann Malik"
date: "April 8, 2021"
output: 
  html_document:
    number_sections: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```
  
The LENGTH ~ NECK model has a residual standard error (the standard deviation of the plotted points from the normal curve) of 5.376 and an adjusted R-squared (how much using a certain variable - in this case NECK - to predict LENGTH improved the prediction than simply using the LENGTH mean model) of 0.7476.
  
## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
  
A bear with a neck of 17 is predicted to have a length of about 52.77.  
A bear with a neck of 22 is predicted to have a length of about 60.99.  
A bear with a neck of 26 is predicted to have a length of about 67.58.  
  
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).
  
### LENGTH ~ AGE 
  
```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)
```
  
The LENGTH ~ AGE model has a residual standard error of 7.511 and an adjusted R-squared of 0.5073. 
  
Compared to the LENGTH ~ NECK model, this one has a higher residual standard error and a lower adjusted R-squared.  
  
### LENGTH ~ HEADLEN
  
```{r}
headlenModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlenModel)
summary.lm(headlenModel)
```
  
The LENGTH ~ HEADLEN model has a residual standard error of 4.235 and an adjusted R-squared of 0.8434. 
  
Compared to the LENGTH ~ NECK model, this one has a lower residual standard error and a higher adjusted R-squared.    
  